Cooling Requirements for the Vertical Shear Instability in Protoplanetary Disks

### 2015-05-08 appendix.tex
new 6.4, merge with youdin's updates
cleaned up figs folder

\subsection{Maximum critical cooling time}\label{max_cool}   
Here we show that, under appropriate conditions, the maximum critical 
cooling time is that for the $M=0$ or fundamental mode. This allows
us to focus on the fundamental mode to obtain an overall cooling
requirement for the VSI.

Consider the simplified dispersion relations for marginal stability,
Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. We write 
$X= \hat{\omega}\khat$, $\theta = (hq\khat)^2$ and  
treat $M$ as a continuous variable. We find from
Eq. \ref{relax_cond_simp1} that  
%\begin{align}
%  &x^4  = x^2 + M(M+1)(1-\theta), \label{simp1}\\
%  &x^2  = \frac{\gamma-1}{h q}\beta_c (1+2M)^2x - 2M(M+1)\label{simp2},
%\end{align}
%where and
\begin{align}
  2 X \frac{d X}{dM} = \frac{X^2(1-\theta)(2M+1)}{X^2 +
  2M(M+1)(1-\theta)}, 
\end{align}
and from Eq. \ref{relax_cond_simp2} that
\begin{align}
  2X\frac{dX}{dM} = \left[X^2 +
  2M(M+1)\right]\left(\frac{d\ln{\beta_c}}{dM} + \frac{4}{2M+1} +
  \frac{d\ln{X}}{dM}\right) - 2(2M+1). 
\end{align}
We eliminate $dX/dM$, making use of Eq. \ref{relax_cond_simp1} in the process, to obtain 
\begin{align}
  (2M+1)\frac{d\ln{\beta_c}}{dM} %= \frac{\mathcal{N}}{\mathcal{D}}, 
  = -\frac{(1-\theta)M(M+1)\left[ 2(2M+1)^2 + 12X^2\right]
  + (3+\theta)X^2}{2 \left[X^2 +
  2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]}.  
\end{align}
%where the denominator $\mathcal{D}$ is
%\begin{align}
%\mathcal{D} = 2 \left[X^2 +
 % 2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]>0, 
%\end{align}
%and the numerator $\mathcal{N}$ is 
%\begin{align}
 % \mathcal{N} = -(1-\theta)\left[ 2M(M+1)(2M+1)^2 + 12X^2M(M+1)\right]
 % - (3+\theta)X^2 < 0.
%\end{align}
Hence,
\begin{align}\label{dbeta_dm}
  \frac{d\beta_c}{dM}<0, 
\end{align}
for all $M\geq 0$ if $\theta\leq 1$, which imply 
$\mathrm{max}(\beta_c)$ occurs at $M=0$. For fixed $\khat$ the
condition $\theta\leq 1$ can be satisfied in sufficiently thin disks.
This means that if $\beta>\beta_c(M=0)\equiv\beta_\mathrm{crit}$, then
all such modes will be stabilized.

>>>>

\subsection{Maximum critical cooling time}\label{max_cool}   
Here we show that, under appropriate conditions, the longest critical 
cooling time is that for the $M=0$ or fundamental mode. This allows
us to focus on the fundamental mode to obtain an overall cooling
requirement for the VSI.

Consider the simplified dispersion relations for marginal stability,
Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2}. We write 
$X= \hat{\omega}\khat$, $\theta = (hq\khat)^2$ and  
treat $M$ as a continuous variable. We find from
Eq. \ref{relax_cond_simp1} that  
%\begin{align}
%  &x^4  = x^2 + M(M+1)(1-\theta), \label{simp1}\\
%  &x^2  = \frac{\gamma-1}{h q}\beta_c (1+2M)^2x - 2M(M+1)\label{simp2},
%\end{align}
%where and
\begin{align}
  2 X \frac{d X}{dM} = \frac{X^2(1-\theta)(2M+1)}{X^2 +
  2M(M+1)(1-\theta)}, 
\end{align}
and from Eq. \ref{relax_cond_simp2} that
\begin{align}
  2X\frac{dX}{dM} = \left[X^2 +
  2M(M+1)\right]\left(\frac{d\ln{\beta_c}}{dM} + \frac{4}{2M+1} +
  \frac{d\ln{X}}{dM}\right) - 2(2M+1). 
\end{align}
We eliminate $dX/dM$, making use of Eq. \ref{relax_cond_simp1} in the process, to obtain 
\begin{align}
  (2M+1)\frac{d\ln{\beta_c}}{dM} %= \frac{\mathcal{N}}{\mathcal{D}}, 
  = -\frac{(1-\theta)M(M+1)\left[ 2(2M+1)^2 + 12X^2\right]
  + (3+\theta)X^2}{2 \left[X^2 +
  2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]}.  
\end{align}
%where the denominator $\mathcal{D}$ is
%\begin{align}
%\mathcal{D} = 2 \left[X^2 +
 % 2M(M+1)(1-\theta)\right]\left[X^2 + 2M(M+1)\right]>0, 
%\end{align}
%and the numerator $\mathcal{N}$ is 
%\begin{align}
 % \mathcal{N} = -(1-\theta)\left[ 2M(M+1)(2M+1)^2 + 12X^2M(M+1)\right]
 % - (3+\theta)X^2 < 0.
%\end{align}
Hence,
\begin{align}\label{dbeta_dm}
  \frac{d\beta_c}{dM}<0, 
\end{align}
for all $M\geq 0$ if $\theta\leq 1$, which imply 
$\mathrm{max}(\beta_c)$ occurs at $M=0$. This result may also be
obtained by explicitly solving 
Eq. \ref{relax_cond_simp1}---\ref{relax_cond_simp2} with $\theta$
treated as a small parameter. For fixed $\khat$ the 
condition $\theta\leq 1$ can be satisfied in sufficiently thin disks.
All all such modes are stabilized if
$\beta>\beta_c(M=0)\equiv\beta_\mathrm{crit}$.

### 2015-05-08 summary.tex
cleaned up figs folder

Perhaps due to these physical differences, CONO operates in different regimes of parameter
space than the VSI.  CONO grows best for longer wavelengths $\khat \gtrsim 1$ and longer 
cooling times $\beta \sim 1$.

@@

\subsubsection{Radiative transfer} 
Our relatively simple treatment of cooling with an idealized dust opacity
could certainly be generalized in future works.  In hotter disks, gas phase opacities
must be considered.   In cold disks there are many choices for a dust opacity which 
varies with grain sizes and compositions. Dust properties could change the viability 
of the VSI for better or worse.  The radiative transfer itself
could be calculated with higher levels of sophistication, as is already being done in
numerical simulations \citep{stoll14}.

>>>>

Perhaps due to these physical differences, CONO operates in different regimes of parameter
space than the VSI.  CONO grows best for longer wavelengths $\khat \lesssim 1$ and longer 
cooling times $\beta \sim 1$.

@@

\subsection{Radiative transfer} 
Our relatively simple treatment of cooling with an idealized dust opacity
could certainly be generalized in future works.  In hotter disks, gas phase opacities
must be considered.   In cold disks there are many choices for a dust opacity which 
varies with grain sizes and compositions. Dust properties could change the viability 
of the VSI for better or worse.  The radiative transfer itself
could be calculated with higher levels of sophistication, as is already being done in
numerical simulations \citep{stoll14}.
